Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections at the LArIAT experiment

### 2018-05-31 Chapter-3/3-TheLArIATExperiment.tex
s

Two ASDQ chips~\cite{ASDQchip} mounted on a mother board plugged into the chamber serve as front end amplifier/discriminator. The chips are connected to a multi-hit TDC~\cite{Sten} which provides a fast OR output used as first level trigger. The TDC time resolution is 1.18~ns/bin and can accept 2 edges per 9~ns.  
The maximum event rate acceptable by the chamber system is 1 MHz: this rate is not a limiting factor considering that \textcolor{red}{the rate of the tertiary particle beam at the first wire chamber is estimated to be less than 15 kHz}. A full spill of data occurring once per supercycle is stored on the TDC board memory at once and read out by a specially designed controller.  We use LVDS cables to carry both power and data between the controller and the TDCs and from the controller to the rest of the DAQ.  
%It is possible to program the time window for acceptance for hits, time offsets, front end threshold, and pulse shaping parameters through the controller via a USB from a PC or through an Ethernet connection.

>>>>

Two ASDQ chips~\cite{ASDQchip} mounted on a mother board plugged into the chamber serve as front end amplifier/discriminator. The chips are connected to a multi-hit TDC~\cite{Sten} which provides a fast OR output used as first level trigger. The TDC time resolution is 1.18~ns/bin and can accept 2 edges per 9~ns.  
The maximum event rate acceptable by the chamber system is 1 MHz.%: this rate is not a limiting factor considering that the rate of the tertiary particle beam at the first wire chamber is estimated to be less than 15 kHz}. 
A full spill of data occurring once per supercycle is stored on the TDC board memory at once and read out by a specially designed controller.  We use LVDS cables to carry both power and data between the controller and the TDCs and from the controller to the rest of the DAQ.  
%It is possible to program the time window for acceptance for hits, time offsets, front end threshold, and pulse shaping parameters through the controller via a USB from a PC or through an Ethernet connection.

### 2018-05-31 Chapter-6/6-PionCrossSection.tex
s

The bottom panel of figure  \ref{fig:FinalXSResPion} shows the relative difference between the measured cross section and the Geant4 predicted cross section.
A couple of considerations are in order here. On one hand, we note the general agreement between the measured and predicted cross sections. We expected the pion cross section in argon to be well modeled in Geant4, given the vast set of measurements available for pion interaction on different targets -- even if no data is available for argon. Thus, this agreement confirms our confidence in the experimental methodology used.
On the other hand, this agreement is not perfect.  As a matter of fact, its imperfections are rather suggestive: we see a slight shift of the delta resonance at higher energies and a small shape difference, with a steeper drop between the delta region and the flatter part at energies greater than 400 MeV. As we showed in Figure \ref{fig:pi1}, the shape of the delta resonance changes as a function of the nuclear target; thus it is plausible that the Geant4 model under estimates the cross section for argon in the delta resonance region and overestimates it in the higher energy region. Since we see room for improvement of the systematic uncertainties, we look forward to seeing if this difference in shape will be confirmed by further analyses, leading to a re-tuning of the model.

>>>>

The bottom panel of figure  \ref{fig:FinalXSResPion} shows the relative deviation between the measured cross section and the Geant4 predicted cross section.
A couple of considerations are in order here. On one hand, we note the general agreement between the measured and predicted cross sections. We expected the pion cross section in argon to be well modeled in Geant4, given the vast set of measurements available for pion interaction on different targets -- even if no previous data is available on argon. Thus, this agreement confirms our confidence in the experimental methodology used.
On the other hand, this agreement is not perfect.  As a matter of fact, its imperfections are rather suggestive: we see a slight shift of the delta resonance at higher energies and a small shape difference, with a steeper drop between the delta region and the flatter part at energies greater than 400 MeV. As we showed in Figure \ref{fig:pi1}, the shape of the delta resonance changes as a function of the nuclear target; thus it is plausible that the Geant4 model under estimates the cross section for argon in the delta resonance region and overestimates it in the higher energy region. Since we see room for improvement of the systematic uncertainties, we look forward to seeing if this difference in shape will be confirmed by further analyses, leading to a re-tuning of the model.

### 2018-05-31 Chapter-7/7-KaonCrossSection.tex
s

The measured cross section is mostly in disagreement with the predicted cross section; in particular it is significantly higher -- on the order of 20\%-30\% higher -- in the bins lower than 400 MeV. The 100 MeV to 400 MeV range is the region of interest for the expected kaons coming from a proton decay event.

\begin{figure}[htb]
\centering
\includegraphics[width=\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
\caption{ \emph{Top:} ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the combined +60A and +100A data samples, statistical uncertainty and systematic uncertainty in black. The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green. \emph{Bottom:} Relative difference between the measured cross section and the Geant4 prediction.} 
\label{fig:FinalXSKaon}
\end{figure}

>>>>

The measured cross section is mostly in disagreement with the predicted cross section; in particular it is significantly higher -- on the order of 20\%-30\% higher -- in the bins lower than 400 MeV. Such difference is unlikely to be accounted for the small proton contamination expected in the data sample. It is worth noticing that the 100 MeV to 400 MeV energy range is the region of interest for the expected kaons coming from a proton decay event. With this result, we will inform a tuning of the Geant4 model for kaon cross section.

\begin{figure}[htb]
\centering
\includegraphics[width=\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
\caption{ \emph{Top:} ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the combined +60A and +100A data samples, statistical uncertainty and systematic uncertainty in black. The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green. \emph{Bottom:} Relative deviation between the measured cross section and the Geant4 prediction.} 
\label{fig:FinalXSKaon}
\end{figure}

### 2018-05-31 Chapter-8/8-Conclusions.tex
s

\chapter{Conclusions}\label{ch:Conclusions}
In the era of neutrino precision physics, of huge liquid argon detectors and of massive amount of information from LArTPCs, a renewed interest in an ancient measurement arises: the measurement of hadronic interactions with matter. With this work, we presented the first ever ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section measurements as a function of the hadron kinetic energy. 
Both the analysis follow a similar workflow and  they rely on beam line detector information as well as both calorimetry and tracking in the LArTPC.

For the ($\pi^-$-Ar) total hadronic cross section measurement, we start by identifying pion beamline candidates through a series of selections on the beamline and TPC information which maximize the number of pions over other particle species (muons and electrons). We use the LArIAT beamline MC to estimate the beam composition of the selected beamline candidates and we propagate them to the LArAIT TPC constructing a properly weighted sample with the DDMC. We apply the thin slice method on the pion candidates and obtain the raw cross section. From the simulated sample, we obtain two corrections accounting for the beamline background contamination and for reconstruction effects. Finally, we apply the corrections to data and measure the cross section.

In order to measure ($K^+$-Ar) total hadronic  argon cross sections, we follow a similar procedure, i.e. we apply the thin slice method on kaon candidates identified in the beamline and obtain the raw cross section. We then apply a MC derived background correction and a correction for detector effects to obtain the cross section. The background correction accounts for the presence of secondary particles in both the interacting and incident histograms and for the presence of decay events in the interacting plot.

The final results for the ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section are shown side by side in the top part of Figure \ref{fig:finalfinal}. The bottom part of the same figure shows the relative deviation from the cross section predicted by Geant4. In the pion case, the measured cross section mostly agrees with the Geant4 prediction. A hint of a difference in shape is present: Geant4 seems to underestimate the cross section in the delta resonance region and overestimate it for energies greater than 400 MeV. Since the   ($\pi^-$-Ar) has never been measured before on Argon and since previous measurements on different nuclei show that the shape of the delta resonance peak varies as a function of the target nucleus, this hint to a shape disagreement is an interesting feature of this measurement which LArIAT will explore with a future development of the analysis.
In the kaon case, the measured cross section is mostly higher than the Geant4 prediction. Contrary to the pion case, previous data on kaons are scarce even on different nuclei; thus, we do not expect the Geant4 prediction to be as solid in the kaon case as it is in the pion case. The nice agreement between the measured cross section and the prediction of the pion case gives us confidence in the methodology, thus confidence in the kaon measurement. With the kaon cross section measurement, we hope to inform the next generation of Geant4 model tuning, thus providing a more reliable tool in the simulation for proton decay searches.

The ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section analyses are the first physics analyses developed by the LArIAT experiment and  will serve as a basis for the future cross section measurements of pions and kaons for the exclusive channels.

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheRealMoneyPlot.pdf}
\includegraphics[width=0.48\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
\caption{\emph{Left:} ($\pi^-$-Ar) total hadronic cross section measurements in the combined sample overlaid with the  Geant4 prediction (green). \emph{Right:} ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured. The respective Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green. } 
\label{fig:finalfinal}
\end{figure}

>>>>

\chapter{Conclusions}\label{ch:Conclusions}
In the era of neutrino precision physics, of huge liquid argon detectors and of massive amount of information from LArTPCs, a renewed interest in an ancient measurement arises: the measurement of hadronic interactions with matter. With this work, we presented the first ever ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section measurements as a function of the hadron kinetic energy. 
Both the analyses follow a similar workflow and  they rely on beam line detector information as well as both calorimetry and tracking in the LArTPC.

For the ($\pi^-$-Ar) total hadronic cross section measurement, we start by identifying pion beamline candidates through a series of selections on the beamline and TPC information which maximize the number of pions over other particle species (muons and electrons). We use the LArIAT beamline MC to estimate the beam composition of the selected beamline candidates and we propagate them to the LArTPC constructing a properly weighted sample with the DDMC. We apply the thin slice method on the pion candidates and obtain the raw cross section. From the simulated sample, we obtain two corrections accounting for the beamline background contamination and for reconstruction effects. Finally, we apply the corrections to data and measure the cross section.

In order to measure ($K^+$-Ar) total hadronic  argon cross sections, we follow a similar procedure, i.e. we apply the thin slice method on kaon candidates identified in the beamline and obtain the raw cross section. We then apply a MC derived background correction and a correction for detector effects to measure the cross section. The background correction accounts for the presence of secondary particles in both the interacting and incident histograms and for the presence of decay events in the interacting plot.

The final results for the ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section are shown side by side in the top part of Figure \ref{fig:finalfinal}. The bottom part of the same figure shows the relative deviation from the cross section predicted by Geant4. In the pion case, the measured cross section mostly agrees with the Geant4 prediction. A hint of a difference in shape is present: Geant4 seems to underestimate the cross section in the delta resonance region and overestimate it for energies greater than 400 MeV. Since the   ($\pi^-$-Ar) has never been measured before on argon and since previous measurements on different nuclei show that the shape of the delta resonance peak varies as a function of the target nucleus, this hint is an interesting feature of the measurement which LArIAT will explore with a future development of the analysis.
In the kaon case, the measured cross section is mostly higher than the Geant4 prediction. Contrary to the pion case, previous measurements of kaon interaction on any nuclei are scarce; thus, we do not expect the Geant4 prediction to be as solid in the kaon case as it is in the pion case -- as illustrated  by Figure \ref{fig:TrueCarbon} for kaon cross section on carbon. The nice agreement between the measured cross section and the prediction of the pion case gives us confidence in the methodology, thus confidence in the kaon measurement. With the kaon cross section measurement, we hope to inform the next generation of Geant4 model tuning, thus providing a more reliable tool in the simulation for proton decay searches.

The ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section analyses are the first physics analyses developed by the LArIAT experiment and  will serve as a basis for the future cross section measurements of pion and kaon cross sections in the exclusive channels.

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheRealMoneyPlot.pdf}
\includegraphics[width=0.48\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
\caption{\emph{Left:} ($\pi^-$-Ar) total hadronic cross section measurement for  scattering angles greater than 5$^\circ$ measured in the combined sample overlaid with the  Geant4 prediction (green). \emph{Right:} ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured. The respective Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green. } 
\label{fig:finalfinal}
\end{figure}

### 2018-05-31 thesis_main.tex
s

\noindent ``\emph{Dunque io ringrazio tutti quanti.} \hfill ``\emph{At last, I thank everyone.}

@@

I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might be like and wanted more.  I came to this vast and sometimes inhospitable land thinking I'd learn a lot of physics (which I did) and hoping I'd keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while). Little did I know that getting my PhD would be first and foremost a journey to self discovery, the beginning of a path to asymptotically tend to the person and the scientist I want to be. A journey impossible without the help of many people, which I will now proceed to thank; let's hope that the acknowledgements won't be longer than the thesis.

Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest with me and for somehow handling this whole US-Italy thing gracefully. I can't believe you still haven't met Bonnie. Ah, and don't worry, I'll always translate things for you.

Bonnie Fleming, my advisor, I don't even know where to start in thanking you. Thanks for listening to me before I proved I was worth be listened to; thanks for always believing I could make it, even at the bottom of my doubts; thanks for all the laughters, for all the freedom, for all the great advices; thanks for leading by example; thanks for being always supportive of all my crazy side projects; thanks for choosing each and every person in the most wonderful research group ever and thanks for letting me be part of it. Thanks for showing me that even Bonnie Freaking Fleming is human, has her downs and still she strives.

@@

Jen Raaf and Jonathan Asaadi, these analyses would have not seen the light of day without you. Thanks for your many inputs and for letting your door open to talk about physics and/or my latest paranoia. Thanks for the patience and understanding. And thanks for leading that small experiment with a big heart called LArIAT and all its crazy ponies. Speaking of which, thanks to all my collaborators, especially Irene Nutini, who started the path for the pion analysis, Greg Pulliam, who suffered with me at the end of said path, Dan the Man, Jason, Johnny, Will Fo and Mr Bojangles. A special token of gratitude goes to Roberto Acciarri and Andrzej M. Szelc, you are so much more to me than collaborators and previous car owners, you know you'll always have a home in Chicago.

Thanks to the MicroBooNE experiment which has been an incredible learning experience. In particular, I want to thank  Sam Zeller for the countless pieces of advice,  my CRT troopers Martin, David, Rui, Igor and Linda  (if I know something about hardware it's because of you), and Kazu and  Kirby (if I know something about software it's because of you).

@@

Shany Danieli and Louise Laage, you're the piece of heart I left in New Haven. Shany, I don't know what made you think I was the cool kid in the group, but I'm so glad you stuck with me. I'm not sure I would have survived my first semester at Yale without you: you shine in your smartness and lift other people with you -- it's a rare quality in our field. Louise, my dear, thanks for always finding the time to listen, for your attuned spirit and for the magnolias.

@@

Finally, thanks to Anto, my husband, my topo and my adventure companion, cause people might think I'm nice and easy, but you know I'm not... And still, you followed me across half the world and you're still here, so that I can never feel alone.

>>>>

\noindent ``\emph{Dunque io ringrazio tutti quanti.} \hfill ``\emph{So, I thank everyone.}

@@

I jumped the pond a little less than 5 years ago because I had a taste of what being a physicists in the US might be like and wanted more.  I came to this vast and sometimes inhospitable land thinking I'd learn a lot of physics (which I did) and hoping I'd keep doing research for as long as I could (which I'm fortunate enough to say I'll still do for a while). Little did I know that getting my PhD would be first and foremost a journey to self discovery, the beginning of a path to asymptotically tend to the person and the scientist I want to be. A journey impossible without the help of many people, which I will now proceed to thank. Let's hope that the acknowledgements won't be longer than the thesis itself.

Mom, you are and will always be the hero of my story; I know this has been challenging for you, thanks for being there for me at every hour of the day and the night, for being always honest and for somehow handling this whole US-Italy thing gracefully. I can't believe you still haven't met Bonnie. Ah, and don't worry, I'll always translate things for you.

Bonnie Fleming, my advisor, I don't even know where to start in thanking you. Thanks for listening to me before I proved I was worth be listened to; thanks for always believing I could make it, even at the bottom of my doubts; thanks for all the laughters, for all the freedom, for all the great advices; thanks for leading by example; thanks for being always supportive of all my crazy side projects; thanks for choosing each and every person in the most wonderful research group ever and thanks for letting me be part of it. Thanks for showing me that even Bonnie %Freaking 
Fleming is human, has her downs and still she thrives.

@@

Jen Raaf and Jonathan Asaadi, these analyses would have not seen the light of day without you. Thanks for your many inputs and for letting your door open to talk about physics and/or my latest paranoia. Thanks for the patience and understanding. And thanks for leading that small experiment with a big heart called LArIAT and all its crazy ponies. Speaking of which, thanks to all my collaborators, especially Irene Nutini, who started the path for the pion analysis, Greg Pulliam, who suffered with me at the end of said path, Dan the Man, Jason, Johnny, Will Fo and Mr Bojangles. A special token of gratitude goes to Roberto Acciarri and Andrzej M. Szelc, you are so much more to me than collaborators and previous car owners, you know you'll always have a home in Chicago (or wherever this job will bring me).

Thanks to the MicroBooNE experiment which has been an incredible learning experience.  I want to thank  Sam Zeller for the countless pieces of advice,  my CRT troopers Martin, David, Rui, Igor and Linda  -- if I know something about hardware it's because of you -- and Kazu and  Kirby, who I shamelessly bothered many times for software help, sushi and coffee.

@@

Shany Danieli, Louise Laage and Flor Orosz Hunziker, you're the piece of heart I left in New Haven. Shany, I don't know what made you think I was the cool kid in the group, but I'm so glad you stuck with me. I'm not sure I would have survived my first semester at Yale without you. Your intelligence is matched only by your kindness: you have the rare gift to shine and make others shine with you. Louise, my dear, thanks for always finding the time to listen, for your attuned spirit and for the magnolias. Flor, thanks for welcoming me to the new world with love and giving me the right definition of ``America".

@@

Finally, thanks to Anto, my husband, my topo and my adventure companion, cause people might think I'm nice and easy, but you know I'm not... And still, you followed me across half the world and you're here, so that I can never feel alone.
